Polytope notes

### 2020-02-26 notes/bo-separated.tex
attempt to re-sketch ideas for linear-separated link proof from discussion with Raf

(Bo, 2020-02-11, attempt to rigorize previous proof sketch)

@@

>>>>

\break
(Bo, 2020-02-11, attempt to rigorize previous proof sketch)

@@

\break
(Bo 2020-02-26, attempt to find constant independent of $p$)

\begin{theorem}
  Suppose $L$ embeds $\ell$ with the $\epsilon$-thickened link.
  Let $L^*$ and $\ell^*$ be the corresponding Bayes risks.
  Then there exists $c > 0$ such that for all $p$ and all $u'$,
  \[ L(u';p) - L^*(p) \geq c \cdot \left[ \ell(\psi(u');p) - \ell^*(p) \right]  . \]
\end{theorem}

This follows from the following theorem:
\begin{theorem}
  Suppose $L$ embeds $\ell$ with the $\epsilon$-thickened link.
  Then for all pairs $r,r' \in \R$, there exists $c > 0$ such that, for all $u \in \psi^{-1}(r)$ and all $u' \in \psi^{-1}(r')$, and for all $p \in \gamma_r$,
  \[ L(u';p) - L(u;p) \geq c \cdot \left[ \ell(r';p) - \ell(r;p) \right] . \]
\end{theorem}
Sketch of ideas as discussed by Raf and Bo last week:

Hope: If we restrict to the relative interior of $\gamma_r$, there is a unique set $U \subseteq \reals^d$ which is $\argmin L(u;p)$ for all $p$ in $relint(\gamma_r)$.

Now, if $u'$ maps to $r'$ then it has distance at least $\epsilon$ from the set $U$.
Let $S$ be the face of $U$ whose relative interior $u'$ projects to.
Recall we wrote $V_S$ for the set of normals to the face and $D_S(p)$ is the set of subgradients of $L$ at that face.
Here we are claiming $V_S$ doesn't depend on $p$ as long as it's in the relative interior, because we hoped above that $U$ is fixed for such $p$; but $D_S(p)$ depends on $p$.
Anyway, we have:
  \[ L(u';p) \geq \epsilon c_S(p) \]
where
  \[ c_S(p) := \inf_{v \in V_S} \sup_{d \in D_S(p)} v \cdot d . \]
Now, since $D_S(p)$ is a polytope, the linear function $v \cdot d$ achieves its maximum at one of the finitely many extreme points.
In fact, $D_S(p)$ is the $p$-weighted Minkowski sum of polytopes $T_S^1,\dots,T_S^n$.
I guess we are arguing that these polytopes are each constant for all of $u$ in the relative interior of the face indexed by $S$.
So there are a finite set of matrixes indexed by $i$, $M_i$, whose $y$th column is an extreme point of $T_S^y$.
So,
  \[ c_S(p) = \inf_{v \in V_S} \max_i v M_i p . \]
Now fix any $i$.
The set of $p$ such that $v M_i p > 0$ is convex.
Furthermore, each $p$ can be associated with at least one such $i$, otherwise, we would get a contradiction that $U$ is not the set we thought it was.
We will use convexity of the set soon.

\vskip1em
Let $p^{\circ}$ be any fixed distribution in the relative interior of $\gamma_r$.
Write any $p$ in $\gamma_r$ as $\lambda p^{\circ} + (1-\lambda) p^b$ for some $p^b$ on the relative boundary.
For each $p^b$, for all small enough $\lambda$, there is a unique $M_i$ associated with all $p$ of this form, thanks to convexity mentioned above.
So for any $p^b$, we get for all associated $p$,
\begin{align*}
  c_S(p) &= \inf_{v \in V_S} v M_i p  \\
         &= \inf_{v \in V_S} \lambda v M_i p^{\circ} + (1-\lambda) v M_i p^{b}  \\
         &\geq \lambda \inf_{v \in V_S} \lambda v M_i p^{\circ}  \\
         &\geq \lambda \alpha_i(p^{\circ})
\end{align*}
for some constant $\alpha_i(p^{\circ}) > 0$, using our previous result/argument (the infimum is achieved by some $v$ for which $v M_i p^{\circ} > 0$, by compactness and continuiity).
(Have to argue about what happened to $p^b$ in the case this doesn't vanish!?)
Let $\alpha(p^{\circ}) = \min_i \alpha_i(p^{\circ})$.

\vskip1em
So we now have that for all $p \in \gamma_r$ parameterized by $\lambda$,
\begin{align*}
  L(u';p) - L(u;p)
    &\geq \epsilon c_S(p)  \\
    &\geq \epsilon \lambda \alpha(p^{\circ}) .
\end{align*}
Meanwhile, writing $v_{rr'} = \ell(r') - \ell(r)$,
\begin{align*}
  \ell(r';p) - \ell(r;p)
    &= v_{rr'} \cdot p  \\
    &= \lambda v_{rr'} \cdot p^{\circ} + (1-\lambda) v_{rr'} \cdot p^{b}  \\
    &\geq \lambda v_{rr'} \cdot p^{\circ}   & \text{(need to justify)}  \\
    &=    \lambda \beta(p^{\circ})
\end{align*}
for some constant $\beta(p^{\circ}) > 0$ (this follows because $p^{\circ}$ is in the relative interior).
Now by choosing $c = \epsilon \alpha(p^{\circ}) / \beta(p^{\circ})$, we prove the theorem.
